\documentclass[11pt]{article}

\usepackage{geometry}
\usepackage{amsmath, amsthm, amssymb, bbm}
\usepackage{hyperref}
\usepackage{setspace}
\usepackage{xcolor}
\usepackage{natbib}
\usepackage{tikz}
\usetikzlibrary{decorations.pathreplacing,calc}

% Define colors for emphasis
\definecolor{mAlert}{RGB}{235, 129, 27}
\newcommand{\alert}[1]{\textcolor{mAlert}{#1}}

\newtheorem{assumption}{Assumption}
\newtheorem{proposition}{Proposition}
\newtheorem{lemma}{Lemma}
\newtheorem{corollary}{Corollary}

\title{AI, Headquarters and Guijie}
\author{Zehao Zhang}
% \date{\today}

\setstretch{1.1}

\begin{document}

\maketitle

\section{Model}

\paragraph{Information.} The state is $\theta\in\{0,1\}$ with prior $\Pr(\theta=1)=p\in(0,1)$. The principal observes a binary signal $s$ with $\Pr(s=1\mid \theta=1)=1$ and $\Pr(s=1\mid \theta=0)=q\in(0,1)$; the agent observes $\theta$ perfectly. 

\paragraph{Timing and contracts.} (i) The principal observes $s$. If $s=1$, she initiates a campaign and offers an outcome-contingent wage scheme $w=(w_0,w_1)$; if $s=0$, no campaign occurs. (ii) The agent observes $\theta$ and $w$ and chooses effort $a\in[0,1)$. (iii) Outcome $y\in\{0,1\}$ realizes with $\Pr(y=1\mid \theta,a)=\theta a$. Effort is unobservable and non-verifiable; only the outcome is contractible. Without loss of generality set $w_0=0$. Both parties are risk neutral.

\paragraph{Technology and payoffs.} The agent's cost is $g(a)$ with $g(0)=0$, $\lim_{a\to 1}g(a)=\infty$, $g'(0)=0$, $g'>0$, $g''>0$. The principal's gross value from success is $r_p>0$. With $w_0=0$, the expected payoffs are
\[
\mathbb{E}[U_P\mid \theta,a]=\theta a\,(r_p-w_1),\qquad \mathbb{E}[U_A\mid \theta,a]=\theta a\,w_1-g(a).
\]
After $s=1$, letting $\pi\equiv \Pr(\theta=1\mid s=1) = \frac{p}{p+(1-p)q}$,
\[
\mathbb{E}[U_P\mid s=1,a]=\pi a\,(r_p-w_1),\qquad \mathbb{E}[U_A\mid s=1,a]=\pi a\,w_1-g(a).
\]

\section{Results}

\paragraph{First-best.} With contractible effort and full information, a planner chooses $a(\theta)$ to maximize $r_p\,\theta\,a-g(a)$. Using $g'(0)=0$ and $g''>0$, the unique solution is $a^{FB}(0)=0$ and, for $\theta=1$, a strictly positive interior effort $a^{FB}\in(0,1)$ defined by $g'(a^{FB})=r_p$ (since $g'(0)=0<r_p$). After $s=1$, expected surplus is $\pi\,[r_p a^{FB}-g(a^{FB})]$.

\paragraph{Second-best.} With hidden effort and outcome-only contracts, the agent's best response to $w_1$ is $a^*(0)=0$ and, for $\theta=1$, $a^*(1)=a$ satisfying $g'(a)=w_1$. The principal equivalently chooses $a$ via $w_1=g'(a)$ to maximize $\pi\,[r_p a-g'(a)\,a]$. The unique optimum $a^{SB}$ is characterized by
\[
r_p = g'(a^{SB}) + a^{SB} g''(a^{SB}),\qquad w_1^{SB}=g'(a^{SB}).
\]
Note that $a^{SB} < a^{FB}$ since $g''>0$.



\end{document}